% $Id$

As a member of the LIGO Scientific Collaboration, I have been fortunate to
have benefited through advice from and discussions with many people. It would
not be possible to thank everyone who I have worked with over the past five
years without making the acknowlegements longest chapter in this dissertation,
so I shall only attempt to thank those who I have interacted with the most and
hope that the others forgive me.

First and foremost, I would like to thank Patrick Brady. I am fortunate to
have an advsior whose as big a drunk as I am.

I would also like to thank Jolien Creighton for his help and enthusiasm over
the past five years. It has been fun working with Jolien and I have learnt a
great deal from him through his patient explanations.

I am greatful to Bruce Allen for suggesting the search for binary inspiral as
a research topic and his assitance with the scientific and computational
obstacles along the way. Thanks also to Gabriela Gonz\'{a}lez for patiently
anserwing my many stupid questions about the LIGO interferometers helping me
understand the data that I have been analyzing.

I would like to thank the members of my committee: Daniel Agterberg, John
Friedman and Leonard Parker for their careful reading of this dissertation and
helpful suggestions for its improvement.

I also would like to thank Warren Anderson, Teviet Creighton, Stephen
Fairhurst, Scott Koranda, Eirini Messaritaki, Ben Owen, Xavier Siemens and
Alan Wiseman for help, advice and pints of beer. I am also indebted to Axel's
for many useful discussions.

Thanks to Steve Nelson, Wyatt Osato and Quiana Robinson for their help in the
preparation of this this and, of course, to Sue Arthur for making everything
run smoothly.

I could not have come this far without the constant love and support of my
parents, to whom this thesis is dedicated. Finally, I would like to thank
Emily Dobbins for all the love and understanding over the past two years.
