%-----------------------------------------------------------------------
%
% File Name: thesis.tex
%
% Author: Brown, D. A.
%
% Revision: $Id$
%
%-----------------------------------------------------------------------

% document class and packages
\documentclass[12pt,notitlepage]{report}
\usepackage{bibunits}
%\usepackage{hyperref}
\usepackage{uwmthesis}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{rotating}
\usepackage{tensor}

\hbadness=10000

% new command definitions
\newcommand{\half}{\frac{1}{2}}
\newcommand{\ospsd}{\ensuremath{S_n\left(\left|f_{k}\right|\right)}}

% journal definitions
\newcommand{\apj}{{\it Astrophysical J.}}
\newcommand{\apjl}{{\it Astrophysical J.}}
\newcommand{\aap}{{\it Astron. and Astrophys.}}
\newcommand{\cmp}{{\it Commun. Math. Phys.}}
\newcommand{\grg}{{\it Gen. Rel. Grav.}}
\newcommand{\cqg}{{\it Class. Quant. Grav.}}
\newcommand{\lr}{{\it Living Reviews in Relativity}}
\newcommand{\mnras}{{\it Mon. Not. Roy. Astr. Soc.}}
\newcommand{\pr}{{\it Phys. Rev.}}
\newcommand{\prl}{{\it Phys. Rev. Lett.}}
\newcommand{\prd}{{\it Phys. Rev. D}}
\newcommand{\prsl}{{\it Proc. R. Soc. Lond. A}}
\newcommand{\ptrsl}{{\it Phil. Trans. Roy. Soc. London}}
\newcommand{\rmp}{{\it Rev. Mod. Phys.}}

\begin{document}
\title{% taken from dissertator fellowship application
Searching for Gravitational Radiation from Black Hole MACHOs 
in the Galactic Halo
}
\author{\bf Duncan A. Brown}
\majorprof{Patrick Brady}
\submitdate{December 2004}
\degree{Doctor of Philosophy}
\program{Physics}
\copyrightyear{2004}
\majordept{Physics}
\havededicationtrue
\dedication{to\\ Mum and Dad}
\haveminorfalse
\copyrighttrue
\doctoratetrue
\figurespagetrue
\tablespagetrue

\Abstract{% taken from dissertator fellowship application
The Laser Interferometer Gravitational Wave Observatory (LIGO) is one of a new
generation of detectors of gravitational radiation. The existence of
gravitational radiation was first predicted by Einstein in 1916, however
gravity waves have not yet been directly observed.

One source of gravitation radiation is binary inspiral. Two compact bodies
orbiting each other, such as a pair of black holes, lose energy to
gravitational radiation. As the system loses energy the bodies spiral towards
each other. This causes their orbital speed and the amount of gravitational
radiation to increase, producing a characteristic ``chirp'' waveform in the
LIGO sensitive band.

In this thesis, matched filtering of LIGO science data is used to search for
low mass binary systems in the halo of dark matter surrounding the Milky Way.
Observations of gravitational microlensing events of stars in the Large
Magellanic Cloud suggest that some fraction of the dark matter in the halo may
be in the form of Massive Astrophysical Compact Halo Objects (MACHOs). It has
been proposed that low mass black holes formed in the early universe may be a
component of the MACHO population; some fraction of these black hole MACHOs
will be in binary systems and detectable by LIGO.

The inspiral from a MACHO binary composed of two 0.5 solar mass black holes
enters the LIGO sensitive band around 40 Hz.  The chirp signal increases in
amplitude and frequency, sweeping through the sensitive band to 4400 Hz in 140
seconds. By using evidence from microlensing events and theoretical
predictions of the population an upper limit is placed on the rate of black
hole MACHO inspirals in the galactic halo.
}
\beforepreface
\prefacesection{Preface}
The work presented in this thesis stems from my participation in the LIGO
Scientific Collaboration. 

\vspace*{0.5cm}
\noindent The upper limit on the rate of binary neutron 
star inspirals quoted in chapter \ref{ch:introduction} is based on

\vspace*{0.25cm}

\noindent B.~Abbott~{\it et~al.} (The LIGO Scientific Collaboration),
``Analysis of LIGO data for gravitational waves from binary neutron stars,''
\prd~{\bf 69} (2004) 122001.

\vspace*{0.5cm}

\noindent Chapter \ref{ch:pipeline} is based on material from

\vspace*{0.25cm}

\noindent Duncan A. Brown~{\it et~al.}, ''Searching for Gravitational Waves
from Binary Inspirals with LIGO,'' accepted for publication in \cqg

\vspace*{0.25cm}

\noindent and

\vspace*{0.25cm}

\noindent B. Abbott~{\it et~al.} (The LIGO Scientific Collaboration), ``Search
for binary neutron star coalescence in the Local Group,'' to be submitted to
\cqg

\vspace*{0.5cm}

\noindent Chapter \ref{ch:hardware} is based on 

\vspace*{0.25cm}

\noindent Duncan A. Brown (for the LIGO Scientific Collaboration), ``Testing
the LIGO inspiral analysis with hardware injections,'' \cqg~{\bf 21}, S797
(2004).

\vspace*{0.5cm}

\noindent Chapter \ref{ch:result} is based on

\vspace*{0.25cm}

\noindent B. Abbott~{\it et~al.} (The LIGO Scientific Collaboration), ``Search
for binary black hole MACHO coalescence in the Galactic Halo,'' to be
submitted to \prd


\prefacesection{Acknowledgments}
\input{acknowledge}

\prefacesection{Conventions}
There are two possible sign conventions for the Fourier transform of a time
domain quantity $v(t)$. In this thesis, we define the Fourier transform
$\tilde{v}(f)$ of a $v(t)$ to be
\begin{equation*}
\tilde{v}(f)=\int_{-\infty}^\infty dt\,v(t)\, e^{- 2 \pi i f t}
\end{equation*}
and the inverse Fourier transform to be 
\begin{equation*}
v(t)=\int_{-\infty}^\infty df\,\tilde{v}(f)\, e^{2 \pi i f t}.
\end{equation*}
This convention differs from that used in some gravitational wave literature,
but is the adopted convention in the LIGO Scientific Collaboration.

\vspace{0.5cm}

\noindent The time-stamps of interferometer data are measured in Global
Positioning System (GPS) seconds: seconds since 00:00.00 UTC January 6, 1980
as measured by an atomic clock.

\vspace{0.5cm}

\noindent Astronomical distances are quoted in parsecs
\begin{equation*}
1\,\mathrm{pc} = 3.0856775807 \times 10^{16}\,\mathrm{m}
\end{equation*}
and masses in units of solar mass
\begin{equation*}
1\,\mathrm{M}_\odot = 1.98892 \times 10^{30}\,\mathrm{kg}.
\end{equation*}

\afterpreface

\Chapter{Introduction}
\label{ch:introduction}
\input{introduction}
\Chapter{Gravitational Radiation from Binary Inspiral}
\label{ch:inspiral}
\input{inspiral}
\Chapter{Binary Black Hole MACHOs}
\label{ch:macho}
\input{macho}
\Chapter{Binary Inspiral Search Algorithms}
\label{ch:findchirp}
\input{findchirp}
\Chapter{Detection Pipelines For Coalescing Binaries}
\label{ch:pipeline}
\input{pipeline}
\Chapter{Hardware Signal Injections}
\label{ch:hardware}
\input{hardware}
%\chapter{Searches for Binary Neutron Star Coalescence}
%\label{ch:bns}
%\input{bns}
\Chapter{The Rate of Binary Black Hole MACHO Inspirals in the Halo}
\label{ch:result}
\input{result}
\Chapter{Conclusion}
\label{ch:conclusion}
\input{conclusion}

\clearpage
\bibliographystyle{unsrt}
\bibliography{references}
\addcontentsline{toc}{chapter}{\numberline {Bibliography}}

\clearpage
\birthplacedate{Nottingham, United Kingdom \>\>January 25, 1976}
\collegewherewhen{%
\>University of Newcastle Upon Tyne \>\>1994--1999, \>M.Math.\\
\>\uwm	\>\>1999--2004, \>Ph.D.}

\begin{startvita}
\end{startvita}

\renewenvironment{thebibliography}[1]%
  {\begin{list}{\labelenumi\hss}%
     {\usecounter{enumi}\setlength{\labelwidth}{3em}%
      \setlength{\leftmargin}{5em}}}%
  {\end{list}}
\renewcommand{\bibitem}[1]{\item\label{#1}\relax}%
\renewcommand{\theenumi}{\arabic{enumi}}%
\begin{publications}
\putbib[papers]
\end{publications}

\finishvita
\end{document}
