There are so many people I would like to thank.
It would not be possible to list every individual that has impacted my life
over the past five years.

First, I would like to thank my advisor, Stefan Ballmer. Your
breadth of knowledge and your intuition in the lab never ceases to amaze me.
You set the standard that I strive to acheive.

I want to thank Duncan for welcoming me into the group 5 years ago and for
making me a better scientist.

To Peter, Thank you for all of your insightful wisdom. I am honored to have
had the opportunity to learn from you.

I would also like to thank all the members of my defense committee --- 
Duncan Brown, Martin Forstner, Seth Jolly, Matt LaHaye, and Peter Saulson ---
%Artuso, John Laiho, Carl Rosenzweig, Peter Saulson, and Jason Wiles ---
for taking time out of their busy schedules to read and critique my work.

From my first days here, I want to thank Andy Lundgren for helping me get
started on learning all the tools and tricks I've relied on to survive.

To the LIGO PSL team, Jan, Benno, Christina, and Patrick, for helping me
understand the PSL and allowing me to invade the PSL code.

Also from my short time at the Livingston site, I must thank Keith Thorne for
getting up and running on everything front-end related. Without your help and
the help from Rolf and Alex I would not have been able to build the digital
system which has been such an integral part of our experiment.

I must also thank Peter Couvares for his help with deep system administration
questions I've needed answers on to get the front-end working.

The machine shop personnel have help immensely through the years. To each of
them, Charlie, Lou, Lester, and Phil I owe much gratitude. From helping me
route the 300' foot long GPS antenna cable through your shop to
providing me with emergency air pressure to float our tables.

To Sally Prasch and Steve Penn, a huge thanks for sharing with me your
expertise in glass work which is embedded in the heart of this experiment.

I want to especially thank my parents for their careful guidance througout my
life. And to my sister for just being there for me throughout my life.

I want to thank my closest family. Your love and support have been invaluable.
To Elizabeth and Henry for all the laughter and joy you have given us.
And last, but not least to my wife Amber: you have helped me in so many ways,
I really cannot imagine getting through this without you and I look forward to
sharing all of the future with you.

%such joy to my life through all of this.


