
%We will soon embark on a new era of discovery.
%A machine which can observe the universe in a whole new way is about to be
%turned on.
%The significance of this cannot be understated.
%We are about to make history.

%Of course, a drammatic introduction wouldn't be complete without a statement
%on how sentient beings have always gazed at the heavens in wonder.
%We still strive for this understanding and must never let it end.
%The struggles of humankind are too great to let our inspirations fizzle.
%Hope for the future is what sustains us.
%
%Something about Gallileo, Newton, Michelson, Einstein...

The physics community has been building gravitational wave detectors for
decades.
The resonant bar detectors of the past have given way to the new
interferometric detectors which have achieved impressive sensitivities.
These detectors are soon to be dramatically surpassed
by detectors with a more sophisticated optical layout, adding
signal recycling to the configuration.

This new generation of gravitational wave detectors are rapidly coming
online.
Advanced LIGO will be at design sensitivity within a year.
These new detectors will see further, by an order of magnitude over
initial LIGO, this corresponds to a 1000 times greater surveilled volume.
%A thousand times more detector volume.
At this sensitivity, one day of Advanced LIGO observations will survey a
larger space-time volume than the 2 years of observations with initial LIGO.
%And with the advancements in lock acquisition, we will surpass the entire
%integrated volume of initial LIGO in the blink of an eye.
aLIGO isn't a simple upgrade, it's literally a new detector.
Every component has been ripped out and replaced.
The laser source is new, pumping out a massive
%1.21 jigaWatts
180 Watts of power.
The radiation pressure noise associated with this power will start to
dominate the displacement noise of the new 40kg test masses.
%it will start to dominate even the new 40kg test masses.
From the state of the art coatings technology to the silicate bonded
monolithic fused silica fiber assemblies we have left virtually
%\footnote{with the interesting exception of the output mode cleaner, which
%serves as the readout point for the gravitational wave signal.}
nothing untouched.

As I am writing this,
the detector in Livingston is already beginning to surpass the
best sensitivity we ever had in initial LIGO. The detector at Hanford will
soon be sealed in it's capsule to embark on a journey into the farthest
reaches of the
universe\footnote{Manufacturing errors in the test mass coatings required a new
set to be installed, delaying the closure of the vacuum system.}.

Reflecting over my time at Syracuse, we have seen the Large Hadron Collider
turn on and confirm the existence of those things we were looking
for\footnote{the Higgs Boson}.
Of course I can't just leave the Higgs Boson as a footnote.
This is what gives matter its mass and as far as we know we can't have
gravitational waves without mass.
Well, we wouldn't exist without mass either, but that's beside the point...

%My time at Syracuse has brought me great joy.
%I have worked with many from throughout LIGO.

%Of course, my greatest joy comes from my family.
%Without them, I'm not sure I would have ever left my office/lab.
%We have had such a great experience here and made many friends of which we
%will never forget.


%For my children, it is your future that I live for.
%May your own inspirations come with great strength.

From this we begin,
\begin{equation}
F = ma\,.
\end{equation}
