\acresetall

\input{fluff}

\section{Gravitational Waves}

The foundation of Einstein's theory of general relativity is that a freely
falling body's motion is governed by the local space-time curvature.
This curvature is in turn influenced by the presence of
matter\footnote{matter being stuff that has mass}.
This matter not only curves the space it occupies, but also curves the space
around it in such a way as to result in an acceleration on an object given by,
\begin{equation}
a = \frac{GM}{r^2} \;.
\end{equation}

Clearly, as matter moves through space, the curvature of space is changing.
%Not only is the curvature changing in the space in which the matter is
%moving, it is also changing away from the mass.
Having established that information cannot travel faster than the speed of light
the 'information' about how space must curve will be delayed.
From the multipole expansion of the mass distribution, the monopole (which is
the first term) is a scalar quantity that is simply the total
mass of the object.
The second term is, called the 'dipole' is a vector which is the sum of all the
bits of mass multiplied by their position vector from a fixed reference point.
The dipole term is identically the center of mass location times the total mass.
This term can change with time, however the first derivative $mv$ (momentum)
is conserved.
The third term is known as the quadrupole term.
It is this term which has a non zero second derivative that gives rise to a
wave equation.
The amplitude of the gravitational wave is,
\begin{equation}
h = \frac{2G}{c^4r}\ddot I \;.
\end{equation}
where the unitless term $h$ is the gravitational wave strain.
This strain is the $\frac{\Delta L}{L}$ perturbation on the background
space-time metric that WE are looking for.
Gravitational waves, being quadrupolar in nature, stretch space-time in one
direction while squeezing it in the orthogonal direction.


We look for the strain perturbations by measuring the distance between two
objects we call test masses.
In the simple case of a two test mass detector (one arm of LIGO) we are only
sensitive to half of the amplitude $h$ (assuming optimal orientation).
The simultaneous stretching and squeezing makes a differential length detector
a natural choice and with two orthogonal arms we are sensitive to the full
amplitude of the wave.
It is important to note that this factor of 2 increase in the signal, though
helpful, is not the primary motivation for two arms. 
The big benefit comes from common noise cancellation.
We can cancel out common length noise in the two arms because the gravitational
wave will couple directly into the differential degree of freedom in the
detector. See figure \ref{fig:ligoschematic}.

\begin{figure}[htb]
\tikzsetnextfilename{ligoschematic}
\centering
\begin{tikzpicture}
%  \draw[thick,rotate=45] (-1,-0.2) rectangle (1,0.2);
  \draw[very thick,red] (-3,0) -- (1.3,0);
  \draw[very thick,red] (0,0) -- (0,1.3);
  \draw[very thick,red] (0,0) -- (0,-2);
  \fill[red] (1.7,-0.15) arc (92:88:116.048) -- ++(0,0.3) arc (272:268:116.048) -- cycle;
  \fill[red] (-0.15,1.7) arc (-2:2:116.048) -- ++(0.3,0) arc (178:182:116.048) -- cycle;
  \draw[ultra thick] (-0.5,-0.5) -- (0.5,0.5);
  \draw[thick] (1.3,-0.5) rectangle (1.7,0.5);
  \draw[thick] (9.8,-0.5) rectangle (10.2,0.5);
  \draw[thick] (-0.5,1.3) rectangle (0.5,1.7);
  \draw[thick] (-0.5,9.8) rectangle (0.5,10.2);
  \draw[ultra thick,blue,<->] (-1,1.3) -- (-1,10.2);
  \draw[ultra thick,dashed,blue,<->] (-1.5,1.3) -- (-1.5,10.2);
  \draw[ultra thick,blue,>-<] (1.3,-1) -- (10.2,-1);
  \draw[ultra thick,dashed,blue,<->] (1.3,-1.5) -- (10.2,-1.5);
\end{tikzpicture}
\caption[LIGO Schematic]{This schematic shows the layout of a basic
    interferometric gravitational wave detector.
    Each arm of the interferometer is a Fabry-Perot cavity which circulates
    the light in the arms, increasing the response of the detector.
    The blue lines indicate the common (dashed) and differential (solid)
    degrees of freedom.
    The Michelson naturally reads out the differential degree of freedom which
    is free of common noise such as the intensity and frequency noise of the
    laser.
    }
\label{fig:ligoschematic}
\end{figure}


%A pea sized amount of neutron star material has the mass of about 100,000
%locomotives.

\section{Angular Control Using Optical Traps}

In Advanced LIGO we use an active feedback control for angular alignment of
the main mirrors.
Angular control is necessary to stabilize the Sidles-Sigg instability
\cite{2006PhLA..354..167S}
which is discussed in more detail in the angular instability section of our
theory paper which is provided in \ref{sec:IV}.

Sensing for this feedback is done with a technique known as wavefront sensing.
Wavefront sensing works by taking the beam reflected from the cavity and
beating it against the reflected sidebands.
Any misalignment results in a 1st order mode beating against the 0 order mode
of the other beam.
This is picked up in a RF quadrant photodiode and demodulated with the local
oscillator for the sidebands placed on the beam producing a \ac{pdh}
style error signal for pitch and yaw.
Since the input test mass and end test mass are at different Guoy phases
their misalignment affects a different linear combination of quadratures of the
beam.
So, in reflection we can place two quadrant photodiodes for sensing at
different Guoy phases.
Then we can transform the four degrees of freedom of the sensor output to the
four degrees of freedom of the mirror alignments.

A couple of important limitations come from this technique. First, we need some
light to make this work. At the light port there is plenty, but at the dark
port we need all of the light we can get for the DC readout where we get the
gravitational wave signal from.
We are shot noise limited here, so every bit of light counts.
The second limitation is that for angular stabilization
we need to control above the hard mode frequency of the instability which,
for Advanced LIGO at high power, is at about 6 Hz.
But we need a sharp cutoff in the feedback below 10Hz in order to not introduce
noise in the sensitive band of LIGO.
There is very little room to get the noise down while keeping the phase of the
feedback right to make the loop stable.


%It is basically a \ac{pdh} lock with quadrant photo-diodes measuring
%the differential \ac{pdh} signal from opposite quadrants of the photo-diode.

