\acresetall

%In this chapter I will cover the theory necessary for the angular optical trap
%which we are striving for and which the Syracuse linear trap is the first major
%step on the road.

%This chapter is mostly excerpts from the paper which provides the theory and
%motivation for the angular optical trap experiment.\cite{Perreca:2014}
%Sections \ref{sec:II}, \ref{sec:III}, \ref{sec:IV}, \ref{sec:V}, \ref{app:A},
%\ref{app:B}, and \ref{app:C} are copied directly from this paper.

%There is sensing noise in any active feedback loop.

We can eliminate the sensing noise from the angular control by replacing the
active feedback with a passive stabilization technique.
For stabilizing the length degree of freedom of an optical cavity, the active
feedback can be reduced or even removed using a passive optical trapping
technique.\cite{Corbitt07}
This technique can be applied to additional degrees of freedom by coupling
multiple length traps through the same mass.
This chapter describes the proposal for optical trapping one angular degree
of freedom using two stable optical springs.

\input{otstab.tex}

\input{otcon.tex}

\input{otanginst.tex}

\input{otradnoise.tex}

%\input{otderive.tex}

\section{Technical Noise}

%
% For demonstrating the angular trap concept we will have several technical
% noise challenges.
%
In addition to the fundamental noises we have discussed, the
experimental demonstration of this angular optical trap concept will
be affected by several technical noise sources.
These noises will affect us in being able to initially acquire lock of the
cavity, the measurement of the optical spring, and ultimately for the optical
spring to remain stable after removing the active feedback.


Since the experiment relies on the force provided by a detuned cavity, the
technical noises which will limit the experiment will ultimately show up as
either mirror position noise or laser frequency noise.
The two will be coupled through the optical spring, but we can think of the
sources independantly for now.

% Laser frequency noise sources
The laser source will have frequency noise of its own. Additionally, there will
be relative frequency noise between the two beams.

% Mirror position noise sources
Mirror position noise will come from seismic and other vibrational noise from the
environment.
Additionally, any amplitude noise in the laser will result in a radiation force
noise on the mirror.

% Performance of the experiment relies on reducing these technical noise sources
Performance of the experiment relies on reducing these technical noise sources.
We will address and attempt to mitigate the laser frequency and intensity noises
by implementing systems similar to a LIGO \ac{psl}.
Seismic and mechanical vibration noises will be mitigated through a suspension
system, and accoustic noise through the use of a vacuum system.





