%-----------------------------------------------------------------------
%
% File Name: thesis.tex
%
% Author: Lough, J. D.
%
%
%-----------------------------------------------------------------------

% document class and packages
\documentclass[12pt,notitlepage]{report}
\usepackage{bibunits}
\usepackage{syrthesis}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{subfig}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{tensor}
\usepackage{acronym}
\usepackage{listings}
\lstset{breaklines=true,
  basicstyle=\ttfamily\footnotesize}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{caption}
%\usepackage{subcaption}
%\usepackage[all]{hypcap}
\usepackage{dsfont}
\usepackage[disable,colorinlistoftodos]{todonotes}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{width=14cm,compat=1.6}
\usetikzlibrary{external,shapes,arrows,calc,positioning}
%\tikzexternalize[prefix=tikz/,mode=list and make]
\tikzexternalize[prefix=tikz/]


% journal definitions
\newcommand{\apj}{{\it Astrophysical J.}}
\newcommand{\apjl}{{\it Astrophysical J.}}
\newcommand{\aap}{{\it Astron. and Astrophys.}}
\newcommand{\cmp}{{\it Commun. Math. Phys.}}
\newcommand{\grg}{{\it Gen. Rel. Grav.}}
\newcommand{\cqg}{{\it Class. Quant. Grav.}}
\newcommand{\lr}{{\it Living Reviews in Relativity}}
\newcommand{\mnras}{{\it Mon. Not. Roy. Astr. Soc.}}
\newcommand{\pr}{{\it Phys. Rev.}}
\newcommand{\prl}{{\it Phys. Rev. Lett.}}
\newcommand{\pra}{{\it Phys. Rev. A}}
\newcommand{\prd}{{\it Phys. Rev. D}}
\newcommand{\nat}{{\it Nature}}
\newcommand{\prsl}{{\it Proc. R. Soc. Lond. A}}
\newcommand{\ptrsl}{{\it Phil. Trans. Roy. Soc. London}}
\newcommand{\rmp}{{\it Rev. Mod. Phys.}}
\newcommand{\jcap}{{\it J. Cosmol. Astropart. Phys.}}

% acronymn definitions
\acrodef{LIGO}{Laser Interferometer Gravitational-wave Observatory}
\acrodef{BNS}{binary neutron star}
\acrodef{BBH}{binary black hole}
\acrodef{NSBH}{neutron-star black-hole binary}
\acrodef{FAR}{\emph{false alarm rate}}
\acrodef{IFAR}{inverse false alarm rate}
\acrodef{SNR}{signal-to-noise ratio}
\acrodef{MM}{minimal match}
\acrodef{IFO}{interferometer}
\acrodef{CBC}{compact binary coalesence}
\acrodef{GW}{gravitational wave}
\acrodef{PDF}{probability distribution function}
\acrodef{S5}{LIGO's fifth science run}
\acrodef{S6}{LIGO's sixth science run}
\acrodef{VSR1}{Virgo's first science run}
\acrodef{VSR2}{Virgo's second science run}
\acrodef{VSR3}{Virgo's third science run}
\acrodef{pN}{post-Newtonian}
\acrodef{LVC}{LIGO Scienctific Collaboration and the Virgo Collaboration}
\acrodef{DAG}{directed acyclic graph}
\acrodef{HIPE}{Hierarchical Inspiral Pipeline Executable}
\acrodef{HIPE}{Hierarchical Inspiral Pipeline Executable}
\acrodef{SPA}{stationary-phase approximation}
\acrodef{ISCO}{the inner-most stable circular orbit}
\acrodef{FFT}{Fast-Fourier Transform}
\acrodef{LLO}{LIGO Livingston Observatory}
\acrodef{LHO}{LIGO Hanford Observatory}
\acrodef{LSC}{LIGO Scientific Collaboration}
\acrodef{DQ}{data quality}
\acrodef{LLF}{local Lorentz frame}
\acrodef{GR}{General Relativity}
\acrodef{EM}{electromagnetic}
\acrodef{pdh}[PDH]{Pound Drever Hall}
\acrodef{psl}[PSL]{pre-stabilized laser}
\acrodef{iss}[ISS]{intensity stabilization servo}
\acrodef{fss}[FSS]{frequency stabilization servo}
\acrodef{pmc}[PMC]{pre-mode cleaner}
\acrodef{scs}[SCS]{Sub-Carrier Servo}
\acrodef{ndyag}[Nd:YAG]{neodymium-doped yttrium aluminum garnet}
\acrodef{npro}[NPRO]{non-planar ring oscillator}
\acrodef{aom}[AOM]{accousto-optic modulator}
\acrodef{eom}[EOM]{electro-optic modulator}
\acrodef{pd}[PD]{photo-diode}
\acrodef{daq}[DAQ]{data acquisitions}
\acrodef{daqd}[daqd]{\ac{daq} daemon}
\acrodef{dtt}[DTT]{Diagnostic Test Tools}
\acrodef{asd}[ASD]{amplitude spectral density}
\acrodef{fsr}[FSR]{free spectral range}
\acrodef{fwhm}[FWHM]{full width at half max}
\acrodef{pzt}[PZT]{piezo electric transducer}
\acrodef{psd}[PSD]{power spectral density}
\acrodef{pbs}[PBS]{polarizing beam splitter}
\acrodef{vco}[VCO]{voltage controlled oscillator}
\acrodef{rin}[RIN]{relative intensity noise}
\acrodef{sos}[SOS]{small optic suspension}
\acrodef{rfpd}[RFPD]{radio frequency photodiode}
\acrodef{fi}[FI]{faraday isolator}
\acrodef{qpd}[QPD]{quadrant photodiode}
\acrodef{rms}[rms]{root mean squared}
\acrodef{osem}[OSEM]{optical sensing electro-magnet}
\acrodef{ligo}[LIGO]{Laser Interferometer Gravitational-wave Observatory}
\acrodef{aligo}[aLIGO]{Advanced \ac{ligo}}
\acrodef{wfs}[WFS]{wavefront sensing}

% common macros
\def\Msun{\ensuremath{\mathrm{M_\odot}}}
\def\Mpc{\ensuremath{\,\mathrm{Mpc}}}
\def\yr{\ensuremath{\,\mathrm{yr}}}
\def\mchirp{\ensuremath{\mathcal{M}}}
\def\mtotal{\ensuremath{\mathrm{M_{total}}}}
\def\d{\ensuremath{\,\mathrm{d}}}
\def\pls{\ensuremath{+}}
\def\crs{\ensuremath{\times}}
\def\effD{\ensuremath{\mathcal{D}}}
\def\th{\ensuremath{^{\mathrm{th}}}}
\def\rhonew{\ensuremath{\rho_{\mathrm{n}}}}
\def\rhonewc{\ensuremath{\rho_{\mathrm{nc}}}}
\def\ihope{\texttt{ihope}}
\def\hipe{\texttt{HIPE}}
\newcommand{\lstin}{\lstinline[basicstyle=\ttfamily]}
\newcommand{\todopar}{\todo[inline,color=blue]}
\newcommand{\todoa}{\todo[color=red]}
\newcommand{\todob}{\todo[color=yellow]}
\newcommand{\todoc}{\todo[color=green]}
\newcommand\KB{1.381e-23}
\newcommand\TROOM{295}
\newcommand\PI{3.1415927}
\newcommand\RHOFS{2.2e3}
\newcommand\EEPOXY{3.378e9}
\newcommand{\tcr}{\textcolor{red}}
\newcommand{\tcb}{\textcolor{blue}}
\newcommand{\tcm}{\textcolor{magenta}}
\newcommand{\tcg}{\textcolor{green}}
\newcommand{\tcp}{\textcolor{purple}}
\newcommand{\irm}{\mathrm{i}}


\begin{document}

\Abstract{

The Advanced LIGO detectors will soon be online with enough sensitivity to
begin detecting gravitational waves,
based on conservative estimates of the rate of
neutron star inspirals.
These first detections are sure to be significant, however, we will always
strive to do better. More questions will be asked about the nature of
neutron star material, rates of black hole inspirals, electromagnetic
counterparts, etc.
To begin to answer all of the questions aLIGO will bring
us we will need even better sensitivity in future gravitational wave
detectors.

This thesis addresses one aspect that will limit us in the future:
angular stability of the test masses.
Angular stability in advanced LIGO uses an active feedback system.
We are proposing to replace the active feedback system with a passive
one, eliminating sensing noise contributions.
This technique uses the radiation pressure of light inside a cavity as a
stable optical spring, fundamentaly the same as technique developed by
Corbitt, et al. \cite{Corbitt07} with an additional degree of freedom.

I will review the theory of the one dimensional technique and discuss the
the multidimensional control theory and angular trap setup.
I will then present results from the one-dimensional trap which we have
built and tested. And propose improvements for the angular trap experiment.

Along the way we have discovered an interesting coupling with thermal
expansion due to round trip absorption in the high reflective coatings.
The front surface HR coating limits our spring stability in this experiment
due to the high circulating power and small beam spot size.


}

\title{
Optical Spring Stabilization
}
\author{James D. Lough}
\pastdegrees{B.S. Virginia Tech, 2001}
\majorprof{Advisor}
\submitdate{December 2014}
\degree{Doctor of Philosophy}
\program{Physics}
\copyrightyear{2014}

\majordept{Physics}
\havededicationtrue
\dedication{to\\my little scientists,\\Elizabeth and Henry}
\haveminorfalse
\copyrighttrue
\doctoratetrue
\figurespagetrue
\tablespagetrue
\signedtitlepfalse

\beforepreface \prefacesection{Preface}

The thesis presented here would not have been possible without collaboration
with my colleagues.
First, and foremost, the main subject of this thesis is the optical trap
experiment taking place at Syracuse University.
Much of the work in setting up the experiment was done together and will
be difficult to differential exact contributions from members.

This experiment was very much a combined effort of primarily four people:
our PI Stefan Ballmer, David Kelley, Antonio Perreca, and myself.
We were fortunate to have some major infrastructure in place when we started:
two large vacuum bell jars, two optics benches with floating legs.
The rest of the experiment was pretty much built from the ground up.

Another graduate student, Fabian Magana-Sandoval recently joined our team
and has contributed to the building of \ac{qpd}s and the commissioning of the
optical lever.

There has also been contributions from other graduate students as part of
requirements for the course Graduate Laboratory:
\begin{itemize}
\item Prayush Kumar designed the \ac{iss}.
\item Alex Nitz worked on the \ac{pmc}.
\item Chris Biwer is adding a feedforward modification to the \ac{pmc}.
\end{itemize}
Also, my work on building and commissioning the \ac{iss} was for fulfillment of the
grad lab course.

My major contributions to the experiment have been:
\begin{itemize}
\item setting up the vacuum system infrastructure and electrical feedthroughs,
\item designing and building the reference cavity and suspension for the
  \ac{fss},
\item designing the trap output mirror "payload" suspension,
\item assembling the digital control system,
\item and assembling and commissioning the suspension control loops.
\end{itemize}

The theory presented in chapter \ref{ch:theory} is primarily copied from our
group's recent paper, Multidimensional optical trapping of a mirror,
Perreca et al. \cite{Perreca:2014}

%The work presented in this thesis stems from my participation in the LIGO
%Scientific Collaboration and the Virgo Collaboration.
%
%A bit of LIGO history and motivation for LIGO itself.
%
%What lies ahead?
%
%Future of radiation pressure applications.

%\prefacesection{Conventions}
%
%Mirror reflectivities are designated by $r$ for the amplitude reflectivity
%and $R$ for the power reflectivity.
%I will use $\mathcal{R}$ for radius of curvature,
%but the distinction should be apparent in the context.



%We use the Einstein summation convention in which repeated upper and lower
%indices are summed over. Greek indices indicate all four spacetime dimensions,
%Latin indicate spatial dimensions only. An over-arrow indicates a vector with
%spatial components only, e.g., $\vec{x} = (x^1,x^2,x^3)$.
%
%\vspace{0.5cm}
%
%\noindent Expectation values of a variable are indicated by $\mathbb{E}()$.
%Averages, unless otherwise noted, are indicated by an overbar.
%
%\vspace{0.5cm}
%
%\noindent The time-stamps of interferometer data are measured in
%Global Positioning System (GPS) seconds: seconds since 00:00.00 UTC
%January 6, 1980 as measured by an atomic clock.

\prefacesection{Acknowledgments}
\input{acknowledge}

\afterpreface

\listoftodos

%\Chapter{Document Planning}
%\label{ch:planning}
%\input{todos.tex}

\Chapter{Introduction}
\label{ch:introduction}
\input{introduction.tex}

%\Chapter{Angular Noise in LIGO Detectors}
%\label{ch:anglenoise}
%\input{angularnoise.tex}

%\Chapter{Optical Spring Stabilization}
%\label{ch:opspringstab}
%\input{optical_spring.tex}

\Chapter{Angular Optical Trap}
\label{ch:theory}
\input{ottheory.tex}

%\Chapter{Angular Trapping}
%\label{ch:angletrap}
%\input{OT_paper.tex}

\Chapter{Pre-Stabilized Laser}
\label{ch:psl}
\input{psl.tex}

\Chapter{Linear Trap Experiment}
\label{ch:lintrapex}
\input{lintrap.tex}

\Chapter{Noise Sources}
\label{ch:noises}
\input{noises.tex}

\Chapter{Results}
\label{ch:results}
\input{results.tex}

\Chapter{Conclusions}
\label{ch:conclusions}
\input{conclusions.tex}

\appendix
\Chapter{Optical Spring Derivation}
\label{ch:otderive}
\input{otderive.tex}

\Chapter{More Derivations}
\label{ch:moreder}
\input{moreder.tex}

\Chapter{Digital System}
\label{ch:digital}
\input{digsys.tex}

%\Chapter{Magic Beans}
%\label{ch:usefulformulas}
%\input{usefulformulas.tex}


%\Chapter{Feedback Control}
%\label{ch:feedback}
%\input(feedback.tex}


\clearpage
\bibliographystyle{unsrt}
\bibliography{OT_paper,references}
\addcontentsline{toc}{chapter}{\numberline {Bibliography}}

\clearpage
\birthplacedate{Blacksburg, Virginia \>\>August 30, 1973}
\collegewherewhen{
\>Virginia Tech \>\>B.S. Physics, 2001}

\newpage
\null%\vskip1in%
\begin{center}
{\Large\bf Vita}
\end{center}

\begin{startvita}
\end{startvita}

\renewenvironment{thebibliography}[1]%
  {\begin{list}{\labelenumi\hss}%
     {\usecounter{enumi}\setlength{\labelwidth}{3em}%
      \setlength{\leftmargin}{5em}}}%
  {\end{list}}
\renewcommand{\bibitem}[1]{\item\label{#1}\relax}%
\renewcommand{\theenumi}{\arabic{enumi}}%
\begin{publications}
\putbib[papers]
\end{publications}


\begin{honorarysocieties}
\end{honorarysocieties}


\end{document}
