I would like to express my sincere gratitude to my advisor, Duncan Brown. I have benefited greatly from his extensive knowledge of gravitational-wave astronomy, and I attribute much of my development as an independent researcher to his mentorship. I am thankful to him for providing many research opportunities and for allowing me the freedom to explore several options to find research that I truly enjoy.

Working in the Syracuse gravitational wave group has been an unforgettable experience. The relaxed and welcoming environment that this group creates and works hard to maintain is what immediately told me I could find a home here, and it's something that I will dearly miss. I am particularly grateful to Peter Saulson for his careful explanations of the LIGO detectors and data, and for always being ready to use his shoe as a pendulum to make a point.

Many thanks to Chris Biwer, Collin Capano, and Alex Nitz. In working with them over the past five years, I learned the fundamentals of how to conduct a gravitational-wave data analysis, and how to develop code for brand new analyses of my own design. I am thankful for all the skills I gained as a result of their expertise.

Thanks to Soumi De, Steven Reyes, and Chaitanya Afle for the innumerable conversations and contributions that have helped me find the solutions to many problems. Thank you for being such amazing research group members.

I would especially like to thank Lindsay DeMarchi. Words cannot express how grateful I am for all the help and support you've given me over the entire span of my graduate school career, from studying for the qual to writing this thesis. Thank you for being the kind of scientist and person that makes the world a better place for everyone.

To all other members of the Syracuse gravitational wave group---Derek Davis, Varun Srivastava, Erik Muniz, Ari Pedersen, Nick Didio, Daniel Vander-Hyde, Amber Lenon---thank you for all the fun and support.

Much of my research would have been impossible without Larne Pekowsky and Eric Sedore. I am thankful for their constant willingness to help with any and all computing questions and requests for resources, and for their infinite patience when I inevitably crashed the cluster.

To Ed Wishnow and Walt Fitelson---I wouldn't be where I am today without the research experience I gained working with you, and without all the great advice you've given me over the years. Thanks for taking a chance on me.

I would like to thank my defense committee members---Duncan Brown, Stefan Ballmer, Eric Coughlin, Scott Watson, Alex Nitz, and William Wylie---for taking the time to review and critique this thesis.

Finally, I would like to thank my parents, to whom this thesis is dedicated. Thank you for being such a great source of inspiration, and for your endless support and belief in me.
