I would like to express my heartfelt thanks to Duncan Brown--I am incredibly lucky to have met an advisor like him. He has been a huge source of inspiration in both my undergraduate and graduate years, as an amazing scientist and an excellent mentor. I have benefited tremendously from his advice as well as his extensive knowledge about gravitational-wave astronomy. Under his guidance I have learned how to become an independent researcher and how to manage my off time when I'm not doing research. I am grateful to have had an advisor like Duncan. 

I am fortunate and grateful to have had the opportunity to collaborate with Alex Nitz on the work presented in Chapters 3--5. His knowledge is invaluable and I am thankful that I was able to learn so much from him.

While I wasn't able to be at Syracuse in person I thank everyone in the gravitational-wave group for making my experience welcoming and engaging. I am thankful to have learned so much about other aspects of the research being conducted in the group. Thank you to Daniel Finstad, Erick Leon, and Soumi De for the help and contributions that have been vital to my research.

To my friends--Sam Usman, Belinda Cheeseboro, Dave Caron, Jaysin Lord, Esraa Ahmad, and Greg Walsh--thank you for all of the support and fun. I would like to thank Sam Usman for the hours long phone calls and amazing advice. I would like to thank Belinda Cheeseboro for being there for me at each road block and for all the laughs.  I would like to thank Dave Caron for letting me spend countless hours in his office working. I would like to thank Jaysin Lord and Greg Walsh for their continuous support and belief in me. I would like to thank Esraa Ahmad for being there for me and comforting me when things got tough. 

Thanks to the West Virginia University friends for all of your support, laughs, and fun chats.
Thank you to Cindy Ramsey, Jim Adams, and the Mountaineer United Soccer Club for your support and for reintroducing me to soccer. I really appreciate the outlet for my stress and the joy that I experience when I play.
Thank you to the VanguardSTEM team for all of your amazing advice, support, encouraging me, and giving me a space for me to be unapologetically me. The community that I gained through VanguardSTEM has continuously inspired me throughout my journey. 

I would like to  thank my defense committee members--Duncan Brown, Stefan Ballmer, Lisa Manning, Jay Hubisz, Alex Nitz, and Natalie Russo--for taking the time to review and critique this dissertation. For my graduate school research, I would thankfully like to acknowledge support from the NSF award AST-1559694 and the Chancellor's Scholars Program. 

Finally, I would like to thank my family--my mom and brothers--to whom this dissertation is dedicated. I thank my brothers, Collin and Jace, and cousin, Zach for our countless hours of DnD, video games, and joking my graduate career to decompress from weeks of hard work.    