The multitude of gravitational-wave detections has created a thrilling era of astrophysics. With each detection and electromagnetic counterpart we have begun to answer many questions about the astrophysical compact objects. Currently binary black holes are routinely detected by the Advanced LIGO and Virgo detectors. The first two detections of binary neutron mergers provided so much insight into their physics, even though they were very different. We now understand that neutron star mergers can produce a gamma-ray bursts and provide us with the heavy elements in our universe as well as allowed us to explore the nuclear equation of state. With improvements to the detectors and the addition of third generation detectors the detection of binary neutron star mergers will also become routine. An increased number of detections will aid us in gaining new insights about the physics of compact objects and how binary these binaries form.

In this thesis, we used observations and modeling to study the eccentricity of binary neutron star systems. Specifically, the work in this thesis focuses on how we can extract information about the binary system through searches, parameter estimation, and future prospects in third generation detectors. We have used the Advanced LIGO and Virgo observations to search for eccentric binary systems and to measure their eccentricity, as well as determine the ability of a third generation detector to detect and measure eccentric binary systems. 

We presented the form of the change in eccentricity and semi-major axis of an eccentric system as well as demonstrating the lifetime for a binary with a given eccentricity. We summarize the methods and algorithms used in the searches for gravitational waves and the measurement of the parameters of the signals.

We conducted a search for eccentric binary neutron stars in Advanced LIGO's first and second observing runs and place an upper limit on their merger. While we did not find any other significant candidates, the limit we have placed aids in the understanding of how well we can detect these mergers with current detectors. We also determine how much data would be needed with A+ and with Cosmic Explorer to make a detection of a binary neutron star merger. 

We provided estimates of the eccentricity of the two binary neutron star mergers detected in the first and second observing runs.
Our analysis measured the eccentricity of GW170817 and GW190425 using a full parameter estimation analysis. We determined that there is a degeneracy between the eccentricity and chirp mass parameters. Previous works done to measure the eccentricity of GW190425 were done~\cite{Romero-Shaw:2020aaj}. While our analysis was not in agreement with theirs, we determined that prior choice and full parameter estimation are important in accurately measuring parameters. The accurate measurement of the parameters is valuable as the eccentricity may provide valuable information about the binary. In the case of binary neutron stars, eccentric systems may have different electromagnetic emission than that of circular systems.

We determined the ability of a third generation detector, Cosmic Explorer, to detect eccentric binary neutron star mergers and to measure their eccentricity. Using template banks and simulations, we found what eccentricities a bank containing circular templates would be effectual at a gravitational-wave frequency of 7~Hz. We showed that a stochastic placement can be used to construct template banks that are effective. We also determined that the computational cost of a matched-filter search for both the quasi-circular and eccentric template banks is not significantly larger than that of Advanced LIGO. We calculated the approximate signal-to-noise ratio needed to distinguish the eccentricity of the binary. 

As Advanced LIGO and Virgo continue to improve in sensitivity there will be more detections of gravitational waves from binary neutron star mergers. As third generation detectors are added to the detector network, the chance of detecting an eccentric gravitational wave from a binary neutron star system increases. With the detection of an eccentric binary neutron star merger we will be able to test the strong-field regime of general relativity, constrain the neutron star equation of state, and ascertain the formation channel of the binary.