Consider a binary system that consists of two neutron stars with mass $m_1$, $m_2$, separated by a distance, d. We can neglect higher order multipole terms and define the gravitational field by the quadrupole formula~\cite{misner:1973} 

\begin{equation}
    h^{TT}_{jk} = \frac{2G}{c^4r}\frac{d^2Q^{TT}_{ij}(t-r)}{dt^2},
\end{equation}

Using conservation laws we can describe the loss of energy and angular momentum from one such system using the quadrupole moment of the binary, $Q_{ij}$. In integral form the conservation laws for energy, E, and angular momentum, L, are given as~\cite{Peters:1964zz, misner:1973}

\begin{equation}
    \begin{split}
        \frac{dE}{dt} &= \int \rho a_i v_i d^3x, \\
        \frac{dL}{dt} &= \int \epsilon_{ijk}x_i\rho a_k d^3x.
    \end{split}
\end{equation}

Since energy and angular momentum are conserved only the reaction part of the potential given by~\cite{misner:1973}
\begin{equation}
    \begin{split}
        &\Phi = \Phi_{Standard Newtonian Theory} + \Phi^{(react)}, \\
        &\Phi^{(react)} = \frac{1}{5}\frac{d^5Q_{ij}}{dt^5}x_ix_j.
    \end{split}
\end{equation}

Eq.~\ref{dedt} states that binaries radiate their eccentricity away over time. The lifetime of these binaries is long depending on the initial eccentricity. As $e \to 1$ the lifetime of the system decreases significantly. For binaries with moderate eccentricity, they will radiate their eccentricity away and can be detected by circular template banks in Advanced LIGO. We will discuss detection of gravitational waves in Sec.~\ref{matched-filtering}.\amber{Maybe make the T vs e plot in Peters 1964}.

The two gravitational wave polarizations states, $h_+$ and $h_\times$, generated from an binary system due to gravitational radiation can be written as~\cite{Damour:2004bz}
\begin{equation}\label{polarizations}
\begin{aligned}
h_{+} &=\frac{1}{2}\left(p_{i} p_{j}-q_{i} q_{j}\right) h_{i j}^{T T}, \\
h_{\times} &=\frac{1}{2}\left(p_{i} q_{j}+p_{j} q_{i}\right) h_{i j}^{T T},
\end{aligned}
\end{equation}
where \textbf{p} and \textbf{q} are two orthogonal unit vectors in the plane of the sky and $h_{i j}^{TT}$ is the transverse traceless component of of the radiation field given by
\begin{equation}
h_{j k}^{\mathrm{TT}}=\frac{2 G}{c^{4} r} \frac{d^{2} \mathcal{Q}_{j k}^{\mathrm{TT}}(t-r)}{d t^{2}}.
\end{equation}

Search pipelines are used to detect gravitational wave signals. In this thesis we will focus on the matched-filtering methods of PyCBC. In Sec.~\ref{matched-filtering}, we describe in detail the algorithms and methods used for detection of gravitational-wave signals in PyCBC. When a detection is made the parameters of the signal are measured using Bayesian parameter estimation. We can use PyCBC Inference~\cite{Biwer:2018osg} to measure the parameters of a gravitational-wave signal, these methods are also described in detail in Sec.~\ref{methods-PE}.

While the Newtonian description is sufficient to describe motion with low velocity it does not accurately describe all motion, like the precession of Mercury, well.

Widely separated binaries lose energy and separation very slowly over billions of years, but the binaries detected by \amber{LIGO and Virgo signify that the separations of these binaries is small enough to be driven to merger through gravitational-wave emission within the age of the universe. This would suggest that the systems observed by LIGO and Virgo are undergoing transitions or formation that bring the components masses of the binary system closer together.} 