I would first like to thank my advisor, Peter Saulson, for his continued mentorship. Your experience, knowledge, 
and passion for teaching have inspired me. You taught me to be excited about my work and to have fun doing it. 
You always make sure that everyone in the room understands the big picture and I hope to carry that philosophy with me. 

The Syracuse University Gravitational Wave Group is a truly unique group of people. Diverse scientific interests, 
expertise, kindness, willingness to teach, and group meetings full of bagels create an environment that fosters 
the development of graduate students and I wouldn't trade my experiences here for any other group in the LSC. 

To Duncan Brown and Stefan Ballmer, thank you for helping me broaden and develop my knowledge of gravitational 
wave astronomy. You have both taught me a tremendous amount and have encouraged all of the members of the Syracuse 
group to diversify their interests and understand the experiment from top to bottom.  

I would like to thank my officemates: Prayush Kumar, Jim Lough, Chris Biwer, Alex Nitz, Soumi De, and Steve Reyes. 
Thank you for all of the companionship and for putting up with my horrible jokes. I expect you to continue them 
once I leave.

To Ryan Fisher, thank you for making sure we enjoyed every minute of our conference travel and for all of the 
advice and mentorship you've provided. 

To Laura Nuttall, thank you for working alongside me, battling through O1 with me, and proofreading so many 
things. Syracuse DetChar was a lot of fun with you here.

To the Detector Characterization group, Josh Smith, Andy Lundgren, Gaby Gonzalez, Peter Saulson, David Shoemaker, 
Jess McIver, Laura Nuttall, Ryan Fisher, Duncan Macleod, Marissa Walker, 
OG Thomas Abbott, Nutsinee Kijbunchoo, Joe Areeda, and everyone else. It's been a pleasure to work with all of you 
for the last 4 years and you've 
all taught me something valuable along the way. I always look forward to the next face-to-face meeting.

I would like to thank my committee members, Howard Blair, Matt LaHaye, Alan Middleton, Duncan Brown, Ryan Fisher, 
and Peter Saulson, for taking time out of their schedules to serve on my thesis defense committee.

To Nathan, Forrest, and Michael, thanks for helping me survive the first few years in Syracuse. It would've been 
infinitely more difficult without you guys. 

To Matt, Danica, Will, Allie, and Joel, thanks for making Syracuse a fun place to live. You are all awesome and 
I hope we don't end up too far away from each other. 

To Mom, Dad, Tim, and Lauren, thank you for always supporting me in everything I've done. 

To all of the Millers and Meszlers, thank you for being a second family to me.  

To Mary, thank you for everything.
