%%%%%%%%%%%%%%%%% Acknowledgements
I've had the privilege over the past several years to work with numerous physicists,
make friends, and build connections through my membership in the LIGO Scientific
collaboration. To those that I fail to thank here, please know that I am appreciative,
and that I intend no slight. 

\noindent I would first like to express my deepest gratitude and thanks to my advisor,
Duncan Brown. His broad knowledge of gravitational-wave astronomy and his
exuberance have been an inspiration. I am indebted to him for his
guidance, encouragement, and especially his patience.

\noindent I would also like to thank the members of my defense committee --- Robert Moucha,
Martin Forstner, Liviu Movilenau, Simon Catterall, and Mitchell Soderberg --- 
for taking time out of their busy schedule to read and critique my work.

\noindent To the Syracuse University gravitational-wave faculty, Peter Saulson 
and Stefan Ballmer, I wold like to thank you both for allowing me to constantly 
pick your brains. Our conversations have been immensely valuable to me, and have
broadened my understanding of the field.

\noindent For his mentorship, when I was a starting graduate student, I would
also like to thank Ian Harry. 

\noindent For always being willing to bounce ideas and putting up with me, I
would like to thank my office mates Chris Biwer and Thomas Massinger.

\noindent I must all thank my parents for their guidance throughout my life.

\noindent I owe a special debt to Keith Riles for introducing me to LIGO,
without whom, I would not have had the amazing experiences and challenges that
come with working on the frontiers of gravitational-wave physics. 
%%%
