%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%% PREFACE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The work presented in this thesis is the result of my participation in the
LIGO Scientific Collaboration (LSC) and does not reflect the
scientific opinion of the LSC.

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:bns_spin} is based on material from

\vspace*{0.25cm}

\noindent   D.~A.~Brown, I.~Harry, A.~Lundgren and A.~H.~Nitz,
``Detecting binary neutron star systems with spin in advanced gravitational-wave detectors,''
{\it Phys. Rev. D} {\bf 86}, 084017 (2012)

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:nsbh_faith} is based on material from

\vspace*{0.25cm}

\noindent    A.~H.~Nitz, A.~Lundgren, D.~A.~Brown, E.~Ochsner, D.~Keppel and I.~W.~Harry,
``Accuracy of gravitational waveform models for observing neutron-star–black-hole binaries in Advanced LIGO,''
{\it Phys. Rev. D} {\bf 88}, no. 12, 124039 (2013)
\vspace*{0.5cm}

\noindent Chapter~\ref{ch:nsbh_prec} is based on material from

\vspace*{0.25cm}

\noindent   I.~W.~Harry, A.~H.~Nitz, D.~A.~Brown, A.~Lundgren, E.~Ochsner and D.~Keppel,
``Investigating the effect of precession on searches for neutron-star-black-hole binaries with Advanced LIGO,''
{\it Phys. Rev. D} {\bf 89}, no. 2, 024010 (2014)

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:single_stage} is based on material from

\vspace*{0.25cm}

\noindent D  S.~A.~Usman {\it et al.}, ``An improved pipeline to search for gravitational waves from compact binary coalescence,''
 arXiv:1508.02357 [gr-qc]. (2015)

\vspace*{0.5cm}

