\thispagestyle{empty}
\begin{center}
{\large {\bf VITA}}
\end{center}
\vspace{1cm}
{\bf AUTHOR}: Alexander Harvey Nitz \\
{\bf NATIONALITY}: American \\
%\vspace{0.5cm}
{\bf DATE OF BIRTH}: December 10, 1987 \\
%\vspace{0.5cm}
{\bf DEGREES AWARDED}:
\begin{itemize}
\item B.A., Physics, University of Michigan, 2010
\end{itemize}
%\vspace{0.5cm}
{\bf PROFESSIONAL EMPLOYMENT}:
\begin{itemize}
\item[] Graduate Teaching/Research Assistant, Department of Physics, Syracuse University (2010-2015).
\end{itemize}
{\bf PUBLICATIONS}:
\begin{itemize}
\item \textit{Accuracy of gravitational waveform models for observing neutron-star–black-hole binaries in Advanced LIGO}, \textbf{Alexander H. Nitz}, Andrew Lundgren, Duncan A. Brown, Evan Ochsner, Drew Keppel, and Ian W. Harry, Phys. Rev. D 88, 124039 (2013)
\item \textit{Template banks to search for low-mass binary black holes in advanced gravitational-wave detectors}, Duncan A. Brown, Prayush Kumar, and \textbf{Alexander H. Nitz}, Phys. Rev. D 87, 082004 (2013)
\item \textit{Detecting binary neutron star systems with spin in advanced gravitational-wave detectors}, Duncan A. Brown, Ian Harry, Andrew Lundgren, and \textbf{Alexander H. Nitz}, Phys. Rev. D 86, 084017 (2012)
\item \textit{Investigating the effect of precession on searches for neutron-star–black-hole binaries with Advanced LIGO}, Ian W. Harry, \textbf{Alexander H. Nitz}, Duncan A. Brown, Andrew P. Lundgren, Evan Ochsner, and Drew Keppel\\, Phys. Rev. D 89, 024010 (2014)
\item \textit{Implementing a search for aligned-spin neutron star-black hole systems with advanced ground based gravitational wave detectors\\, Dal Canton, et al., Phys. Rev. D 90, 082004 (2014)

\end{itemize}
{\bf AWARDS}:
\begin{itemize}
\item 2015 Levinstein Fellowship for Outstanding Graduate Student
\item Graduate Assistantship, Department of Physics, Syracuse University (2010-2015).
\item Dean's Merit Scholarship, University of Michigan, USA (2006).
\end{itemize}
{\bf TEACHING}:
\begin{itemize}
\item ASTRO 101 (Fall 2010/2011).
\end{itemize}
{\bf PRESENTATIONS}:
\begin{itemize}
\item Searching for Gravitational Waves from Spinning Binary Neutron Star Sources, Eastern Gravity Meeting, Syracuse University, Nov 2012
\item Faithfulness of Gravitational Wave Models for Spinning NSBH Sources, Eastern Gravity Meeting, University of Toronto, June 2013 \\
\item Constructing CBC Searches with PyCBC, LSC-Virgo Meeting, Stanford University, September 2014
\end{itemize}
