Once this thesis exceeded 256 pages, my advisor challenged me to write 512. At
first I did not think that would be possible. But as I sit here contemplating
all the people that have my appreciation, I realize I could easily reach that
number, and exceed it. Still, for the sake of brevity, and the members of my
committee, I must limit my acknowledgements to a few individuals. I hope anyone
I do not mention here will forgive me.

First, I would like to thank my advisor, Duncan Brown. Had he not asked me what
I was planning to do three years ago for a research topic, I would not be here
today. His mentorship, and his friendship, have made this all possible.

I would also like to thank the other members of my defense committee --- Peter
Saulson, Stefan Ballmer, Britton Plourde, Alan Middleton, and Scott Sampson ---
for taking time out of their busy schedules to read and critique my work.

I have learned much from all of the members of the LIGO Scientific
Collaboration and the Virgo Collaboration. I would particularly like to thank
Ian Harry, Dave McKechan, Thomas Dent, Duncan Macleod, Steve Fairhurst,
Stuncan, Alan Weinstein, Drew Keppel, and Kipp Cannon for all their help, and
for collaborating with me on a number of projects. A special thanks also goes
to Andrew Lundgren for knowing, well, everything.

Closer to home, I would like to thank the members of the Syracuse University
Gravitational Wave Group, particularly Peter Couvares, for all of the
stimulating discussions over coffee and noodle bowls. A special thanks goes to
Matt West and Larne Pekowsky for entertaining my hair-brained ideas. I could
not have asked for better office mates. I would also like to thank Larne for
answering all of my UNIX questions, and for the shared experience of writing
our dissertations.

Outside of the gravity group, I would like to thank David Niedzwieki and Don
Bunk for all the laughs, awkward drawings, and unicorns. I am particularly
indebted to the Vanderpools --- Scot, Jo, and Connor --- for opening up their
home to me these past several weeks, and for all the free meals and games of
Carcassonne they have shared with me the last three years. Thanks also to Jill
Sneider, Janet Epstein, Kelly Sprinkle, and the men of AMI for giving me a
sense of purpose and a diversion away from physics. Summers will never be as
fun without playing softball with the members of the Bad News Bohrs; I shall
miss them all terribly.

My life would not be the same without Kate Vanderpool. Her love and
encouragement have made the last three years pass in a blink, and I cannot thank
her enough.

I would like to thank my step-mother, Teresa Capano, for her guidance. A
special thanks goes to my late Uncle Philly; without his financial support
attending college would have been a challenge. And of course, I am eternally
grateful to my father, for giving me an appreciation for all Nature's
mysteries, and my mother, for giving me the drive to see my studies through.

Finally, I would like to thank Syracuse University and the SU Physics
Department for giving me the opportunity to undertake both graduate, and
undergraduate studies at their institution. I came to Syracuse almost ten
years ago to the day, in August of 2001. In that time I have met and befriended
many wonderful people, some of whom have gone on to bigger and better things.
It feels odd to be the one leaving Syracuse now, but I will always hold the
city, the university, and the people I have met here close to my heart.
