%-----------------------------------------------------------------------
%
% File Name: thesis.tex
%
% Author: Capano, C. D.
%
%
%-----------------------------------------------------------------------

% document class and packages
\documentclass[12pt,notitlepage]{report}
\usepackage{bibunits}
\usepackage{gradschool_thesis}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{tensor}
\usepackage{acronym}
\usepackage{listings}
\lstset{breaklines=true, basicstyle=\ttfamily}
\usepackage{lscape}
\usepackage{hyperref}

% journal definitions
\newcommand{\apj}{{\it Astrophysical J.}}
\newcommand{\apjl}{{\it Astrophysical J.}}
\newcommand{\aap}{{\it Astron. and Astrophys.}}
\newcommand{\cmp}{{\it Commun. Math. Phys.}}
\newcommand{\grg}{{\it Gen. Rel. Grav.}}
\newcommand{\cqg}{{\it Class. Quant. Grav.}}
\newcommand{\lr}{{\it Living Reviews in Relativity}}
\newcommand{\mnras}{{\it Mon. Not. Roy. Astr. Soc.}}
\newcommand{\pr}{{\it Phys. Rev.}}
\newcommand{\prl}{{\it Phys. Rev. Lett.}}
\newcommand{\prd}{{\it Phys. Rev. D}}
\newcommand{\prsl}{{\it Proc. R. Soc. Lond. A}}
\newcommand{\ptrsl}{{\it Phil. Trans. Roy. Soc. London}}
\newcommand{\rmp}{{\it Rev. Mod. Phys.}}

% acronymn definitions
\acrodef{LIGO}{Laser Interferometer Gravitational-wave Observatory}
\acrodef{BNS}{binary neutron star}
\acrodef{BBH}{binary black hole}
\acrodef{NSBH}{neutron-star black-hole binary}
\acrodef{FAR}{\emph{false alarm rate}}
\acrodef{IFAR}{inverse false alarm rate}
\acrodef{SNR}{signal-to-noise ratio}
\acrodef{MM}{minimal match}
\acrodef{IFO}{interferometer}
\acrodef{CBC}{compact binary coalesence}
\acrodef{GW}{gravitational wave}
\acrodef{PDF}{probability distribution function}
\acrodef{PSD}{power spectral density}
\acrodef{S5}{LIGO's fifth science run}
\acrodef{S6}{LIGO's sixth science run}
\acrodef{VSR1}{Virgo's first science run}
\acrodef{VSR2}{Virgo's second science run}
\acrodef{VSR3}{Virgo's third science run}
\acrodef{pN}{post-Newtonian}
\acrodef{LVC}{LIGO Scienctific Collaboration and the Virgo Collaboration}
\acrodef{DAG}{directed acyclic graph}
\acrodef{HIPE}{Hierarchical Inspiral Pipeline Executable}
\acrodef{HIPE}{Hierarchical Inspiral Pipeline Executable}
\acrodef{SPA}{stationary-phase approximation}
\acrodef{ISCO}{the inner-most stable circular orbit}
\acrodef{FFT}{Fast-Fourier Transform}
\acrodef{LLO}{LIGO Livingston Observatory}
\acrodef{LHO}{LIGO Hanford Observatory}
\acrodef{LSC}{LIGO Scientific Collaboration}
\acrodef{DQ}{data quality}
\acrodef{LLF}{local Lorentz frame}
\acrodef{GR}{General Relativity}
\acrodef{EM}{electromagnetic}

% common macros
\def\Msun{\ensuremath{\mathrm{M_\odot}}}
\def\Mpc{\ensuremath{\,\mathrm{Mpc}}}
\def\yr{\ensuremath{\,\mathrm{yr}}}
\def\mchirp{\ensuremath{\mathcal{M}}}
\def\mtotal{\ensuremath{\mathrm{M_{total}}}}
\def\d{\ensuremath{\,\mathrm{d}}}
\def\pls{\ensuremath{+}}
\def\crs{\ensuremath{\times}}
\def\effD{\ensuremath{\mathcal{D}}}
\def\th{\ensuremath{^{\mathrm{th}}}}
\def\rhonew{\ensuremath{\rho_{\mathrm{n}}}}
\def\rhonewc{\ensuremath{\rho_{\mathrm{nc}}}}
\def\ihope{\texttt{ihope}}
\def\hipe{\texttt{HIPE}}

\begin{document}

% The grad school requires the first page to be blank
\newpage
\thispagestyle{empty}
\mbox{}
\newpage

\Abstract{

This thesis describes current efforts to search for gravitational waves from
compact binary coalescences (CBCs) by the LIGO Scientific Collaboration (LSC)
and the Virgo Collaboration. We briefly review the physics of gravitational
wave emission and detection, describing how they are emitted from
``inspiraling" compact stellar mass objects and how the LSC and Virgo try to
detect them using interferometers. Next we review the data-analysis principles
used to search for potential signals in the detectors' noise. These principles
are employed by ``\ihope," which is the data-analysis pipeline used to search
for CBCs. We describe each step in this pipeline and discuss how interferometer
data is stored and examined. Next we present the results from a six-month long
search which occurred in early 2007, during LIGO's fifth science run. This is
followed by details of tuning studies carried out on LIGO's sixth-, and Virgo's
second- and third-, science runs (S6, VSR2, and VSR3), which ran from July 2009
to October 2010. No gravitational waves were detected in these searches. A
``blind injection" was performed during S6/VSR3 and detected by our pipeline,
however.  We detail studies into assigning a statistical significance to this
injection.  Next we use these studies to show that we can expect to detect
gravitational waves with high significance using two detectors in the advanced
detector era.  Finally, we review some future developements for the CBC
pipeline currently being undertaken.

}

\title{% taken from dissertator fellowship application
Searching for Gravitational Waves from Compact Binary Coalescence using LIGO and Virgo Data
}
\author{\bf Collin D. Capano}
\majorprof{Duncan A. Brown}
\submitdate{August 2011}
\degree{Doctor of Philosophy}
\program{Physics}
\copyrightyear{2011}

\majordept{Physics}
\havededicationtrue
\dedication{to\\ my Mom and my Dad}
\haveminorfalse
\copyrighttrue
\doctoratetrue
\figurespagetrue
\tablespagetrue

\beforepreface \prefacesection{Preface}

The work presented in this thesis stems from my participation in the LIGO
Scientific Collaboration and the Virgo Collaboration.

\vspace*{0.5cm}

\noindent Chapter \ref{ch:s5_results} presents results described in:

\vspace*{0.25cm}
\noindent B. P. Abbott et al., ``Search for Gravitational Waves from Low Mass Compact Binary Coalescence in 186 Days of LIGO's fifth Science Run," {\it Phys. Rev.,} {\bf D80}, 047101 (2009)

\vspace*{0.25cm}
\noindent The author was the primary analyst on one month of this data, and contributed to the results presented therein.

\vspace*{0.5cm}
\noindent Chapter \ref{ch:s6_results} presents some results described in:

\vspace*{0.25cm}
\noindent J. Abadie et al., ``Search for Gravitational Waves from Low Mass Compact Binary Coalesence in LIGO's Sixth Science Run and Virgo's Science Runs 2 and 3," {\it In preparation}.

\vspace*{0.25cm}
\noindent As the results discussed in Chapter \ref{ch:s6_results} are currently under review, they are potentially subject to change. Additionally, some of the plots presented in the chapter --- namely those relating to data quality investigations --- will not be published and so have not been subject to the same rigorous review that published plots and results undergo. They therefore do not represent the scientific opinion of the LIGO Scientific Collaboration or the Virgo Collaboration.

\prefacesection{Conventions}

We use the Einstein summation convention in which repeated upper and lower
indices are summed over. Greek indices indicate all four spacetime dimensions,
Latin indicate spatial dimensions only. An over-arrow indicates a vector with
spatial components only, e.g., $\vec{x} = (x^1,x^2,x^3)$.

\vspace{0.5cm}

\noindent Expectation values of a variable are indicated by $\mathbb{E}()$.
Averages, unless otherwise noted, are indicated by an overbar.

\vspace{0.5cm}

\noindent The time-stamps of interferometer data are measured in
Global Positioning System (GPS) seconds: seconds since 00:00.00 UTC
January 6, 1980 as measured by an atomic clock.

\prefacesection{Acknowledgments}
\input{acknowledge}

\afterpreface

\Chapter{Introduction}
\label{ch:introduction}
\input{introduction}

\Chapter{Physics of Gravitational Waves and the LIGO/Virgo Interferometers}
\label{ch:theory}
\input{theory}

\Chapter{Obtaining Gravitational Wave Triggers from Interferometer Data}
\label{ch:pipeline_principles}
\input{pipeline_principles}

\Chapter{Ranking Triggers by False Alarm Rate}
\label{ch:far}
\input{far}

\Chapter{The IHOPE Pipeline}
\label{ch:ihope_pipeline}
\input{ihope_pipeline}

\Chapter{S5 Results}
\label{ch:s5_results}
\input{s5_results}

\Chapter{S6 Tuning and Results}
\label{ch:s6_results}
\input{s6_results}

\Chapter{Can We Detect with Two Detectors? A Study for LIGO South}
\label{ch:ligo_south_study}
\input{ligo_south_study}

\Chapter{Future Developments and Conclusions}
\label{ch:future_developments}
\input{future_developments}

\clearpage
\bibliographystyle{unsrt}
\bibliography{references}
\addcontentsline{toc}{chapter}{\numberline {Bibliography}}

\clearpage
\birthplacedate{Glens Falls, New York \>\>February 15, 1983}
\collegewherewhen{
\>Syracuse University \>\>B.S. Physics, 2005\\
\>Syracuse University \>\>M.S. Physics, 2008}

\newpage
\null%\vskip1in%
\begin{center}
{\Large\bf Curriculum Vitae}
\end{center}
%\vskip 2em
%\begin{tabbing}
%\tabset
%Title of Dissertation\\
%\>Searching for Gravitational Radiation from Compact Binary
%Coalescence Using LIGO and Virgo Data
%\end{tabbing}
%\vskip 1em

\begin{startvita}
\end{startvita}

\renewenvironment{thebibliography}[1]%
  {\begin{list}{\labelenumi\hss}%
     {\usecounter{enumi}\setlength{\labelwidth}{3em}%
      \setlength{\leftmargin}{5em}}}%
  {\end{list}}
\renewcommand{\bibitem}[1]{\item\label{#1}\relax}%
\renewcommand{\theenumi}{\arabic{enumi}}%
\begin{publications}
\putbib[papers]
\end{publications}

\begin{honorarysocieties}
2001 \>Syracuse University Founders' Scholarship\\
\end{honorarysocieties}

%\finishvita

% The grad school requires the last page to be blank
\newpage
\thispagestyle{empty}
\mbox{}
\end{document}
