I would like to first thank my advisor, Duncan Brown, for his support and mentorship throughout my PhD. Learning
to scientifically is a challenging endeavor. I am grateful for Duncan's guidance through
the entire process. His experience and suggestions have been a constant help in attaining my research goals.

I am grateful to the entire Syracuse University Gravitational wave group for their unique perspectives and for their
diverse expertise. Science, I believe, is best done in the context of diversity, community, and passion, and I believe
the group deeply embodies this ethic in a way that few places could rival.

I am grateful to Peter Saulson and Stefan Ballmer for their helpful perspectives, critiques, and advice in the
progress and development of the research I have conducted here at Syracuse University. I
have grown significantly in my knowledge and confidence in the field of
gravitaitonal wave astronomy due to their help.

I would like to thank my coworkers and officemates: Thomas Vo, Jaysin Lord, TJ Massinger, Soumi De, Chris Biwer,
Lorena Maga$\tilde{\text{n}}$a Zertuche, Swetha Bhagwat, Daniel Finstad, Varun Srivastava,
Chaitanya Afle, Derek Davis, Daniel Vander-Hyde, Nick Didio,
Fabian Maga$\tilde{\text{n}}$a-Sandoval, Ari Pederson, Eric Mu$\tilde{\text{n}}$iz, Gaby , and my virtual-officemate Alex Urban.
Thank you for your friendship and constant encouragement.
Graduate school wouldn't have been nearly as enjoyable, enlightening, or even possible without all of you.
I hope each of you continues to
cultivate the same culture of inclusiveness and openness that you have helped foster here at Syracuse University
or wherever you end up. You all have been great fun to be around.

To the PyCBC group, of which there are too many to list, I thank you for your guidance, encouragement, and
expertise. I have benefited enormously from being in the presence of world-class scientists who have helped
me in every which way to complete this PhD. Thank you. 

I would like to thank my committee members, A, B, C, D, E, 
and Duncan Brown, for taking time out of their schedules to serve on my thesis defense committee.

To the other (under)graduate students at Syracuse University who have befriended me over the past years, I would
like to thank you all as well. I hope you all continue to be excellent. 

To my parents, my siblings, my cousins, my aunts, my uncles, and my entire family, you've given me every opportunity
possible to become anything that I could ever wish to be. Thank you for everything.
