%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%%%%%%%%%% PREFACE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The work presented in this thesis stems from my participation in the
LIGO Scientific Collaboration (LSC) and the NINJA-2 Collaboration.  This
work does not reflect the scientific opinion of the LSC and it was not
reviewed by the collaboration.

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:EOBF2_Effectualness} is based on material from

\vspace*{0.25cm}

\noindent D. A. Brown, P. Kumar and A. H. Nitz, ``Template banks to 
search for low-mass binary black holes with Advanced Gravitational-wave 
detectors,''
{\it Phys. Rev. D} {\bf 87}, 082004 (2013)

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:NRHyb_bank} is based on material from

\vspace*{0.25cm}

\noindent P. Kumar, I. MacDonald, D. A. Brown, H. P. Pfeiffer, K. Cannon,
M. Boyle, L. E. Kidder, A. H. Mroue, B. Szilagyi and A. Zeninoglu, 
``Template banks for binary black hole searches with numerical relativity
waveforms,''
{\it Phys. Rev. D} {\bf 89}, 042002 (2014)

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:ninja2} is based on material from

\vspace*{0.25cm}

\noindent The LIGO Scientific Collaboration, the Virgo Collaboration, 
the NINJA-2 Collaboration, ``The NINJA-2 project: 
Detecting and characterizing gravitational waveforms modelled using 
numerical binary black hole simulations,''
{\it arXiv:1401.0939} (2014) {\rm Submitted to Classical and Quantum Gravity}

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:insSFIMRI} is based on material from

\vspace*{0.25cm}

\noindent E. A. Huerta, P. Kumar, D. A. Brown, 
``Accurate modeling of intermediate-mass-ratio inspirals:
Exploring the form of the self-force in the intermediate-mass-ratio regime,''
{\it Phys. Rev. D} {\bf 86}, 024024 (2012)

\vspace*{0.5cm}

\noindent Chapter~\ref{ch:imrSFIMRI} is based on material from

\vspace*{0.25cm}

\noindent E. A. Huerta, P. Kumar, J. R. Gair, S. T. McWilliams, 
``Self-forced evolutions of an implicit rotating source:
a natural framework to model comparable and intermediate mass-ratio 
systems from inspiral through ringdown,''
{\it arXiv:1403.0561} (2014) {\rm Submitted to Phys. Rev. D}

\vspace*{0.5cm}
