%Will be added before submission to the graduate school.
\iffalse
Duncan \\
Stefan and Peter \\
Jim \\
Morgan \\
Daniel Siegel \\
Thank Enrico \& Ilya for their excellent mentorship on my work as a fellow at the Kavli Summer Program and the work that continued after the program. \\
Larne My work has heavily relied on \\
TJ \\
Chris \\
Collin Capano has been a constant collaborator in all my PyCBC projects. I have relied on him for reviewing almost every piece of code I have committed to PyCBC, and am grateful for the skills I have developed working with him. \\%to him for his guidance and collaboration. \\
I am also very thankful to Alex Nitz for answering many questions in several of my data analysis projects. \\
Daniel, Steven, Swetha, \& Amber \\ 
Laura and Ben \\
other members : Daniel Vander-Hyde, Chaitanya, Derek, Fabian, Lorena, Varun, Ari \\
other collaborators: Ingo, Andrea, Rosa, Edo, \\

sister, mom and dad \\

Special thanks to KITP, mention programs, acknowledgements from those - DARK
Mention grants and fellowships
\fi

I would first like to thank Duncan Brown---I have been incredibly lucky to be mentored by an advisor like him. He has been a constant source of inspiration and support in the past five years, and I have benefited immensely from his extensive knowledge and expertise in gravitational-wave astronomy. 
%I am incredibly lucky to have Duncan Brown as my advisor, who has been a constant source of inspiration and support in the past five years. I have benefited immensely from his extensive knowledge and expertise in gravitational-wave astronomy. 
Working with Duncan, I have learnt how any problem, regardless of its complexity level, can be understood and communicated effectively in simplified steps, and I hope to never forget this in the future. I am exceedingly grateful to him for his invaluable guidance and advice in all aspects of research, for giving me the freedom to explore a variety of research topics, and for numerous exciting research opportunities---starting with participation in the analysis of the first gravitational-wave observations, during my second week in graduate school.

I have thoroughly enjoyed working in the Syracuse gravitational wave group. The informal, welcoming, and engaging environment encouraged in this group, made graduate school an experience that I will miss. I am grateful to Peter Saulson and Stefan Ballmer for their many valuable insights and advice on my research projects, and for useful discussions related to the LIGO detectors and detections at group meetings.

I am fortunate to have had the opportunity to collaborate with Jim Lattimer on analyses of gravitational-wave observations from neutron star mergers, and the work presented in Chapter 3 is a product of one of such projects. I am thankful for getting to learn so much about neutron star physics from him.

It has been a great pleasure working with Morgan Macleod, who very nicely introduced me to hydrodynamics. Chapter 4 is a product of the project that I worked on in close collaboration with him. I am thankful to him for teaching me so much about common envelope physics, and for patiently helping me understand several aspects of the problem.

To Enrico Ramirez-Ruiz and Ilya Mandel, for their mentorship on my project as a fellow at the Kavli summer program in astrophysics, 2017, and on the work that continued after the program.

It has been wonderful working with Daniel Siegel, and my work with him is presented in Chapter 5. I am thankful to him for kindly sharing so much knowledge that helped in broadening my understanding of electromagnetic counterpart aspects of neutron star mergers.

Many thanks to Collin Capano, Chris Biwer, and Alex Nitz. A significant portion of my research in the past five years was conducted in collaboration with them, on many gravitational-wave data analysis and PyCBC development projects. I have learnt a great deal working with them, and I am grateful for all those skills.

I would like to thank T. J. Massinger, who helped me a lot, and got me up to speed with several aspects of research, when I started graduate school during LIGO's first observing run.

Progress on much of my research would have been difficult without Larne Pekowsky. I am indebted to him for helping me cross many computing hurdles, and for willing to answer any kind of computational questions, in the kindest way possible. I would also like to thank Eric Sedore, for promptly helping with computing resources, whenever I reached out.

To Daniel Finstad, Steven Reyes, Swetha Bhagwat, Amber Lenon, and Chaitanya Afle, whose help and contributions have been vital in finding solutions to many problems. Thank you for being brilliant research group members.

To Laura Nuttall and Ben Lackey---thank you for all your help. To all other members of the Syracuse gravitational wave group---Daniel Vander-Hyde, Derek Davis, Fabian Maga\~{n}a-Sandoval, Lorena Maga\~{n}a Zertuche, Varun Srivastava, Ari Pedersen---thank you for all the fun and support.

To Rosa Wallace Everson, Andrea Antoni, Ingo Tews, and Edo Berger---thank you for all your insights and crucial contributions that helped in extracting valuable results from the projects that we collaborated on.

To Sukanta Bose---had I not got the chance to work with him back in India, I probably would not have been here. I am grateful to him for the research experience I gained, and for his excellent mentorship that significantly influenced my decision to apply to graduate schools in the USA.

I would like to thank my defense committee members---Duncan Brown, Stefan Ballmer, Mitchell Soderberg, Scott Watson, Eric Coughlin, and Laura-Anne Minkoff-Zern---for taking the time to review this dissertation.

For my graduate school research, I would like to gratefully acknowledge support from the NSF awards PHY-1707954 and PHY-1607169, the Inaugural Kathy '73 and Stan '72 Walters Endowed Fund for Science Research graduate fellowship, and the Research Excellence Doctoral Funding fellowship. A major part of the work presented here originated at several workshops and conferences, for which I which I would like to acknowledge the following. Work in Chapter 3 was stimulated by the rapid-response workshop ``Astrophysics from a Neutron Star Merger, 2017'' held at the Kavli Institute for Theoretical Physics (KITP). Work presented in Chapter 4 originated at the ``Kavli Summer Program in Astrophysics, 2017'' supported by the Kavli Foundation and the Danish National Research Foundation, with hospitality provided by the Niels Bohr Institute. Work in Chapter 5 originated at the KITP program ``The New Era of Gravitational-Wave Physics and Astrophysics, 2019''. I thank the organizers of the programs mentioned above, and the respective institutes, for providing these wonderful opportunities for research and collaboration. %to produce exciting research ideas.

Finally, I would like to thank my family---my parents and sister, to whom this thesis is dedicated. I thank my sister, Reshmi, for being a great source of strength, and my parents for their care, support, and time over the years.



