Gravitational waves are one of the most remarkable predictions of
Einstein's general theory of relativity~\cite{Einstein:1916}.  In a
sense they are an immediate consequence of reconciling gravity with
special relativity.  As nothing can travel faster than light,
information about the change in position of an object can not
propagate instantaneously.  The gravitational field therefore becomes
a dynamical entity, much like the electromagnetic field in Maxwell's
equations.

There is currently compelling evidence evidence for the existence of
gravitational waves.  These waves carry energy away from their source,
and the energy in a gravitationally-bound system of two stars or black
holes correlates directly to the separation between them.  It is
therefore possible to track the change in separation of such a binary
system, infer the power being radiated away, and compare this to the
predicted power in gravitational waves.  Hulse and Taylor were awarded
the Nobel prize in 1994 for the discovery of a system exhibiting such
behavior~\cite{Hulse:1994, Taylor:1994}. Continued observation of this
system over the intervening years have shown that the loss of energy
matches the prediction from general relativity to within
0.2\%~\cite{Weisberg:2004hi}.


Direct detection of gravitational waves would do far more than further
confirm the prediction of their existence.  Such detections would open
a new window on the universe.  Radio, X-ray and microwave astronomy
have each provided views of the Universe not available though the
others, and each has taught us a great deal.  The gravitational
spectrum is a completely new, and so far unexplored, spectrum.  The
gravitational spectrum can probe regions that the electromagnetic
spectrum can not, including black holes, the interiors of neutron
stars and supernovae, and the very early universe.

The attempts to directly detect gravitational waves are hampered by
the fact that they interact very weakly with matter.  This is a
consequence of the fact that gravity is a very weak force; a small
magnet can, over a short distance, balance the gravitational pull of
the entire Earth.  Nevertheless, direct detection is possible.  There
is currently a worldwide effort underway using interferometric
techniques

\begin{itemize}

\item The Laser Interferometer Gravitational wave Observatory (LIGO)
in the US, with detectors in Hanford, WA and Livingston LA.

\item Virgo, in Cascina, Italy.

\item GEO 600, near Sarstedt, Germany

\item The Large-scale Cryogenic Gravitational wave Telescope (LCGT) is
presently about to being construction in Japan.
\end{itemize}

The LIGO, Virgo and GEO detectors have all reached their initial
design sensitivities enabling detection of gravitational waves from
the final stages of systems like the Hulse-Taylor binary out to (in
the case of LIGO) as far as 45 Megaparsecs (Mpc).  As a point of
reference, the Virgo supercluster extends about 33 Mpc and contains
thousands of galaxies, including our own.

Although there are many potential sources of gravitational waves
a particularly promising one is the inspiral and coalescence of
binary systems containing compact objects, neutron stars or black
holes~\cite{thorne.k:1987} (collectively called \emph{compact binary
coalescence}, henceforth CBC).  Over a large range of masses the
frequencies at which such systems emit the most power are the
frequencies at which LIGO and Virgo are the most sensitive, from about
40 Hz to 1000 Hz.

Despite this remarkable sensitivity, signals in the LIGO data will
likely be quiet relative to the noise levels, Sophisticated data
analysis techniques will be required to extract the signals from noise
and infer properties of their sources.

In addition to the achievements at LIGO and other sites, recent years
have also seen major progress in the field of \emph{numerical
relativity}, the use of computers to accurately simulate systems
governed by general relativity.  From Pretorius' first successful
simulation of the final orbit and merger of two black holes in
2005~\cite{Pretorius:2005gq}, there are now routine simulations of
binary black hole systems with a wide variety of parameters, extending
through several orbits.  There are also an increasing number of
simulations of systems containing matter, including both black holes
in gaseous environments and neutron stars and supernovae.  

The goal of the research presented in this dissertation is to improve
the efficiencies of CBC searches.  In order to claim a detection a
candidate signal must stand out sufficiently above the noise.
Efficiency can therefore be improved either by increasing the
significance of the signal or by reducing the noise.  We take both
approaches here.

The remainder of this dissertation is organized as follows:

\begin{itemize}

\item In chapter~\ref{ch:theory} we present the theory of
gravitational waves, starting from general relativity and deriving
the effect of waves on matter.  We also discuss both analytic and
computational approaches to predicting the form of gravitational
waves.

\item In chapter~\ref{ch:ligo_detectors} we start from the effect of
gravitational waves on matter to discuss the principles of
interferometric detectors. 

\item In chapter~\ref{ch:search} we describe the search for CBC
signals in data from interferometric detectors.

\item Chapter~\ref{comparison} begins the process of utilizing
predictions from numerical relativity to improve the efficiency of
searches.  We proceed by comparing the analytic models used in
searches to a high-accuracy simulation from the Caltech-Cornell group.
Based on these comparisons we derive a number of possible improvements
to the search,

\item Chapter~\ref{ninja1} introduces the first Numerical Injection
Analysis (NINJA) project, a collaboration between numerical
relativists and gravitational-wave astronomers.  The goal of NINJA is
to study the effectiveness of numerous search methods at detecting
gravitational waves and extracting information about their sources.
We apply the recommendations from the previous chapter to the NINJA
data, which simulates the output from LIGO and contains signals from
numerical relativity.

\item Chapter~\ref{ch:ninja2} discusses some of the limitations of the
first NINJA project and introduces NINJA2.  A notable feature of
NINJA2 is the requirements placed on the NR submissions.  We discuss 
the comprehensive analyses that have been performed to ensure that the
submissions meet these requirements.

\item Chapter~\ref{ch:ninja2_results} discusses the construction of the
simulated data sets for NINJA2 and presents preliminary results from
the CBC analysis of the data.  We note some unexpected results that
need further study and open questions that we plan to address in
NINJA2.

\item In chapter~\ref{ch:detchar} we change focus to the reduction of
detector noise.  We proceed by constructing a reduced, simplified
version of the CBC search.  This was used in the latest runs of the
LIGO and Virgo detectors to identify times of excess noise in the
detector and remove these times from analysis.  We include some open
questions about the implementation details of the search revealed
during these analyses.


\end{itemize}

